Russian sanctions figures

### 2024-03-14 ms-russia-figures.Rmd
change panel labels

consumption <- consumption %>%
  # Filter consumption to when Russia is involved
  filter(source_country_iso3c == "RUS"|exporter_iso3c == "RUS"|consumer_iso3c == "RUS") %>%
  # Add variables for whether source/importer/exporter imposed sanctions following annexation of Crimea
    mutate(sanctions_towards_russia = case_when(
      consumer_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", 
                            "NLD", # FIXIT: Need to find out what is happening with the Netherlands and others directly importing
                            "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Imposed sanctions",
      TRUE ~ "Did not impose sanctions"
    )) %>%
      mutate(exporter_sanctions_towards_russia = case_when(
      exporter_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Imposed sanctions",
      TRUE ~ "Did not impose sanctions")) %>%
    mutate(source_sanctions_towards_russia = case_when(
      source_country_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Imposed sanctions",
      TRUE ~ "Did not impose sanctions")) %>% 
  # Add common names (for both sciname and sciname_hs_modified)
    left_join(sciname_metadata %>% 
              select(sciname, common_name), 
            by = "sciname") %>%
  left_join(sciname_metadata %>% 
              select(sciname, "common_name_hs_modified" = "common_name"), 
            by = c("sciname_hs_modified" = "sciname"))

@@

```{r}
# Consumption of Russian products by sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Fill version
consumption %>%
  filter(consumer_iso3c != "RUS", 
         source_country_iso3c == "RUS") %>%
  mutate(path = case_when(
    exporter_iso3c == "RUS" ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, fill = path)) +
  geom_area() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_fill_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Consumption of Russian seafood \n(mil t, live weight)", fill = "") +
  facet_wrap(~sanctions_towards_russia) +
  theme_minimal() +
  theme(legend.position = "bottom") # FIXIT: year labels shouldn't overlap

# Consumption of Russian products by sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Line version
consumption %>%
  filter(consumer_iso3c != "RUS", 
         source_country_iso3c == "RUS") %>%
  mutate(path = case_when(
    exporter_iso3c == "RUS" ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, color = path)) +
  geom_line() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_color_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Consumption of Russian seafood \n(mil t, live weight)", color = "") +
  facet_wrap(~sanctions_towards_russia) +
  theme_minimal() +
  theme(legend.position = "bottom") # FIXIT: year labels shouldn't overlap

# Change in consumption stats
# Change in total consumption of Russian products - imposed sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Imposed sanctions", 
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

@@

# Change in consumption of Russian products via direct import - imposed sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Imposed sanctions", 
         exporter_iso3c == "RUS",
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

@@

# Change in consumption of Russian products via indirect import - imposed sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Imposed sanctions", 
         exporter_iso3c != "RUS",
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

@@

# Change in total consumption of Russian products  - didn't impose sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Did not impose sanctions", 
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

@@

```{r}
# Russian consumption of seafood products from sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Fill version
consumption %>%
  filter(consumer_iso3c == "RUS", 
         source_country_iso3c != "RUS") %>%
  mutate(path = case_when(
    source_country_iso3c == exporter_iso3c ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, source_sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, fill = path)) +
  geom_area() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_fill_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Russian consumption of foreign seafood \n(mil t, live weight)", fill = "") +
  facet_wrap(~source_sanctions_towards_russia) +
  theme_minimal() +
  theme(legend.position = "bottom") # FIXIT: year labels shouldn't overlap

# Russian consumption of seafood products from sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Line version
consumption %>%
  filter(consumer_iso3c == "RUS", 
         source_country_iso3c != "RUS") %>%
  mutate(path = case_when(
    source_country_iso3c == exporter_iso3c ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, source_sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, color = path)) +
  geom_line() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_color_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Russian consumption of foreign seafood \n(mil t, live weight)", color = "") +
  facet_wrap(~source_sanctions_towards_russia) +
  theme_minimal() +
  theme(legend.position = "bottom") # FIXIT: year labels shouldn't overlap

# Change in Russia's consumption from countries imposing sanctions
consumption_change <- consumption %>%
  filter(source_sanctions_towards_russia == "Imposed sanctions", 
         consumer_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

@@

# Change in Russia's consumption from countries imposing sanctions - direct imports
consumption_change <- consumption %>%
  filter(source_sanctions_towards_russia == "Imposed sanctions", 
         source_country_iso3c == exporter_iso3c,
         consumer_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

@@

>>>>

consumption <- consumption %>%
  # Filter consumption to when Russia is involved
  filter(source_country_iso3c == "RUS"|exporter_iso3c == "RUS"|consumer_iso3c == "RUS") %>%
  # Add variables for whether source/importer/exporter imposed sanctions following annexation of Crimea
    mutate(sanctions_towards_russia = case_when(
      consumer_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", 
                            "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Russia unfriendly countries",
      TRUE ~ "Russia friendly countries"
    )) %>%
      mutate(exporter_sanctions_towards_russia = case_when(
      exporter_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Russia unfriendly countries",
      TRUE ~ "Russia friendly countries")) %>%
    mutate(source_sanctions_towards_russia = case_when(
      source_country_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Russia unfriendly countries",
      TRUE ~ "Russia friendly countries")) %>% 
  # Add common names (for both sciname and sciname_hs_modified)
    left_join(sciname_metadata %>% 
              select(sciname, common_name), 
            by = "sciname") %>%
  left_join(sciname_metadata %>% 
              select(sciname, "common_name_hs_modified" = "common_name"), 
            by = c("sciname_hs_modified" = "sciname"))

@@

```{r}
# Consumption of Russian products by sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Fill version
consumption_russian_ts_fill <- consumption %>%
  filter(consumer_iso3c != "RUS", 
         source_country_iso3c == "RUS") %>%
  mutate(path = case_when(
    exporter_iso3c == "RUS" ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, fill = path)) +
  geom_area() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_fill_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Consumption of Russian seafood \n(mil t, live weight)", fill = "") +
  facet_wrap(~sanctions_towards_russia) +
  theme_minimal() +
  theme(legend.position = "bottom") # FIXIT: year labels shouldn't overlap

consumption_russian_ts_fill

# Consumption of Russian products by sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Line version
consumption_russian_ts_line <- consumption %>%
  filter(consumer_iso3c != "RUS", 
         source_country_iso3c == "RUS") %>%
  mutate(path = case_when(
    exporter_iso3c == "RUS" ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, color = path)) +
  geom_line() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_color_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Consumption of Russian seafood \n(mil t, live weight)", color = "") +
  facet_wrap(~sanctions_towards_russia) +
  theme_minimal() +
  theme(legend.position = "bottom") # FIXIT: year labels shouldn't overlap

consumption_russian_ts_line

# Change in consumption stats
# Change in total consumption of Russian products - imposed sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Russia unfriendly countries", 
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

@@

# Change in consumption of Russian products via direct import - imposed sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Russia unfriendly countries", 
         exporter_iso3c == "RUS",
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

@@

# Change in consumption of Russian products via indirect import - imposed sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Russia unfriendly countries", 
         exporter_iso3c != "RUS",
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

@@

# Change in total consumption of Russian products  - didn't impose sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Russia friendly countries", 
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

@@

```{r}
# Russian consumption of seafood products from sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)

# Fill version
consumption_by_russia_ts_fill <- consumption %>%
  filter(consumer_iso3c == "RUS", 
         source_country_iso3c != "RUS") %>%
  mutate(path = case_when(
    source_country_iso3c == exporter_iso3c ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, source_sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, fill = path)) +
  geom_area() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_fill_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Russian consumption of foreign seafood \n(mil t, live weight)", fill = "") +
  facet_wrap(~source_sanctions_towards_russia) +
  theme_minimal() +
  theme(legend.position = "bottom")

consumption_by_russia_ts_fill

# Russian consumption of seafood products from sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Line version
consumption_by_russia_ts_line <- consumption %>%
  filter(consumer_iso3c == "RUS", 
         source_country_iso3c != "RUS") %>%
  mutate(path = case_when(
    source_country_iso3c == exporter_iso3c ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, source_sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, color = path)) +
  geom_line() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_color_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Russian consumption of foreign seafood \n(mil t, live weight)", color = "") +
  facet_wrap(~source_sanctions_towards_russia) +
  theme_minimal() +
  theme(legend.position = "bottom") # FIXIT: year labels shouldn't overlap

consumption_by_russia_ts_line

# Change in Russia's consumption from countries imposing sanctions
consumption_change <- consumption %>%
  filter(source_sanctions_towards_russia == "Russia unfriendly countries", 
         consumer_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

@@

# Change in Russia's consumption from countries imposing sanctions - direct imports
consumption_change <- consumption %>%
  filter(source_sanctions_towards_russia == "Russia unfriendly countries", 
         source_country_iso3c == exporter_iso3c,
         consumer_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

@@

```
```{r, fig.height=6}
# Join plots
ggarrange(consumption_russian_ts_fill, consumption_by_russia_ts_fill, 
          common.legend = TRUE, ncol = 1)

ggarrange(consumption_russian_ts_line, consumption_by_russia_ts_line, 
          common.legend = TRUE, ncol = 1)
```
